---
title: "4-assignment-esm-262"
author: "Emma Rasmussen"
format: html
editor: visual
---

# Building your first function

### Defining contract

This function calculates Simpson's Index.

*Input:* Number of individuals per species and total individuals of all species

*Output:* Simpson's Index (D), value between 0 to 1

The value represents the probability that 2 individuals randomly selected from a sample will be the same species.

*Parameters:*

species, the number of individuals per species

total, the total of all individuals across species

*Return:*

index, the output of Simpson's

*Function:* The function will be the sum of each individual species total ("species") divided by the total of all species ("total") squared.

D = ∑((species /total)^2^

```{r}
#' @param species individuals per species
#' @param total sum of all individuals across all species
#' @return index (D)
```

Function Definition

```{r}

# function definition
index = function(species, total = sum(species)) {
  # here's were I calculate power
  result = sum((species/total)^2)
  return(result)
}
```

Testing function with example data

```{r}
# generate test data frame
test_data_s1 <- data_frame(
  species_name = c("redleg","arroyo","yellowleg","catree","bajachorus","spadefoot","sierranchorus","bullfrog","catoad","torrent","newt","giantsal"),
  individuals = sample(1:50, size = 12)
)

test_data_s2 <- data_frame(
  species_name = c("redleg","arroyo","yellowleg","catree","bajachorus","spadefoot","sierranchorus","bullfrog","catoad","torrent","newt","giantsal"),
  individuals = sample(0:20, size = 12)
)


# run data in function 

site1 <- index(species = test_data_s1$individuals)

site2 <- index(species = test_data_s2$individuals)


```
