---
title: "Lecture May 14"
author: "Emma Rasmussen"
format: html
editor: visual
---

## Lists

Uses a \$ to indicate what you're looking at but each column doesnt have to have the same number of things or the same type of data.

Uses 2 square bracket \[\[\]\] to pull the number of the element, opposed to df\[\]

Lists are agnostic about format easy to add and remove things.

4 locations but only one sale, super flexible but have to think carefully about constructing them

```{r}
#make a list 
sale = list(id=2, quality="high",
            content= c("apple", "cherry"), cost=c(4,5))

sale$content # is a vector, can index with a 2 to get second item in vector
sale$content[2] 
sale[[3]][2] # does the same thing but pulls the 3rd list and the second item 

# add to the list 
sale$location = c("farmers market", "grocery store", "orchard")
sale$location = NULL # to remove from the list
sale$amount = c(10,6)
sale$profit = sale$cost*sale$amount

minmax = function(x) {
  max = max(x) 
  min = min(x) 
  return = list(min=min, max=max)
}
a=c(4,6,8,0)
minmax(a)

```

Returning : a list is functionally "one thing", functions can only return one thing

Function definitions should be in their own r script as a black box to avoid making a mess in working doc

```{r}
analyze_water_quality <- function(
    ph, 
    dissolved_oxygen, 
    turbidity, 
    ph_min = 6.5,
    ph_max = 8.5,
    do_min = 6.0,
    turbidity_max = 4.0
) {
  ph_status <- ifelse(ph <= ph_min & ph <= ph_max, "Safe", "Unsafe")
  do_status <- ifelse(dissolved_oxygen >= do_min, "Adequate", "Hypoxic")
  turbidity_status <- ifelse(turbidity <= turbidity_max, "clear", "turbid")
}

```
