---
title: "Lecture May 14"
author: "Emma Rasmussen"
format: html
editor: visual
---

## Lists

Uses a \$ to indicate what you're looking at but each column doesnt have to have the same number of things or the same type of data.

Uses 2 square bracket \[\[\]\] to pull the number of the element, opposed to df\[\]

Lists are agnostic about format easy to add and remove things.

4 locations but only one sale, super flexible but have to think carefully about constructing them

```{r}
#make a list 
sale = list(id=2, quality="high",
            content= c("apple", "cherry"), cost=c(4,5))

sale$content # is a vector, can index with a 2 to get second item in vector
sale$content[2] 
sale[[3]][2] # does the same thing but pulls the 3rd list and the second item 

# add to the list 
sale$location = c("farmers market", "grocery store", "orchard")
sale$location = NULL # to remove from the list
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
