---
title: "lecture6b"
author: "Emma Rasmussen"
format: html
editor: visual
---

## Flow Control

Classic example is IF then ELSE

ifelse(condition, TRUE, FALSE)

best used for simple application

if..else structure

```{r}
a = 20
classr  = ifelse(a > 90, "high","low")
classr

# also works with multiple values
a = runif(min=1, max=100, n=5)
a

# ifelse can work with vectors
classr  = ifelse(a > 90, "high","low")
classr

```

Flow Control summarize_diversity2

```{r}
library(here)
source(here("esm-262/summarize_diversity2.R")) # i dont have this 

flowers = sample(c("rose","peony","daisy"), replace=T, prob = c(0.1, 0.8, 0.1),size=30)

summarize_diversity(flowers, index="B")$simpson
```

Ifelse good for short simple conditions,

complex: if {} else{}

but multiple conditions:

use case_when (requires tidyverse) or switch

case_when uses \~ to indicate condition

```{r}
library(tidyverse)
a = runif(min = 1, max = 60, n=5)
classr = ifelse((a>10), ifelse(a>50, "high", "med"), "low")
classr


# using case_when
case_when((fish == "steelhead") ~ "A", (fish == "tuna")~ "B", !(fish == "steelhead") & !(fish == "tuna") ~ "C")

#make a function with case_when
classify_fish = function(fish) {
  class = case_when((fish == "steelhead") ~ "A", (fish == "tuna")~ "B", !(fish == "steelhead") & !(fish == "tuna") ~ "C") 
  return(class) }

 classify_fish("tuna")


```

Design a function

Inputs:

-   nutrient concentration (\@ least 5 days)

-   temperature for at least 5 days

-   thresholds for temperature and nutrient (with default values)

Outputs:

-   mean nutrient concentration

-   pollution risk as "low", "med" or "high"

```{r}
library(here)
here("esm-262/classify_lake.R")

# generate some data to try the function
temperature = runif(min=4, max=35, n=30)
nutrient = runif(min=5, max=40, n=30)

# try it
res=classify_lake(temperature=temperature, nutrient=nutrient)

# lets pick some numbers we know should give us a low value
nutrient = runif(min=0, max=5, n=30)
classify_lake(temperature=temperature, nutrient=nutrient)


# now try high
# use repeat to generate the same value multiple times
nutrient = rep(50, times=30)
temperature = rep(25, times=30)
classify_lake(temperature=temperature, nutrient=nutrient)

##### Error checking 
power_gen = function(height, flow, rho=1000, g=9.8, Keff=0.8) {
  # make sure inputs are positive and if not set as NA so result will be NA
   if (height < 0) height=NA
  
 # an alterative is to exit the function prematurely 
  if (flow < 0) return("Flow cannot be less than 0")
  #    return("Height cannot be less than 0")
      
  # calculate power
    result = rho * height * flow * g * Keff
  return(result)
}

#power_gen(2,3)
#power_gen(-2,3)
power_gen(2,-3)
```

Error Checking: what will function do
