---
title: "lecture6_API"
format: html
editor: visual
---

## API = Application Programming Interface

Many APIs need a key – unique identifier that: tells server who you are, lets providers track & limit usage, usually free for students/researchers

DO NOT put key into the code keep private by: storing as a variable

```{r}
# in your .Renviron file (run usethis::edit_r_environ() to open it):
# OPENAQ_KEY=your_key_here

# then in your R script, retrieve it safely:
# my_key <- Sys.getenv("OPENAQ_KEY")
```

Workflow:

-   Find the API – read the documentation to understand endpoints and parameters

-   Build a request – use request and req_url_query, functions to construct your query

-   Send it – req_perform() to get the response

-   Parse the response – usually resp_body_json() to turn JSON into an R doesnt always work resp_content_type() can help

## Packages Needed

```{r}
#install.packages("httr2")
#install.packages("jsonlite")
library(jsonlite)
library(httr2)
library(tidyverse)
```

## Practice

Open-Meteo Air Quality API - free, no key needed

```{r}
# Build the request 
req <- request ("https://air-quality-api.open-meteo.com/v1/air-quality") %>% 
  req_url_query(
    latitude = 34.42, 
    longitude = -119.70, 
    hourly = "pm2_5", 
    past_days = 3 # last 3 days of data 
  )

# send the quest and parse the response

resp <- req %>% req_perform()

# Convert to an R list 
aq_raw <- resp %>% resp_body_json(simplifyVector = TRUE) #keeps NAs

# Show what we got back 
names(aq_raw)
str(aq_raw) #compact overview of structure 
length(aq_raw$pm2_5) #number of observations


```

Translate JSON R list to an R data frame

```{r}

# Extract what you need for the df 
aq_df <- data.frame(
  time = as.POSIXct(unlist(aq_raw$hourly$time), format = "%Y-%m-%dT%H:%M"), # shows how to interpret JSON date format 
  pm2_5 = unlist(aq_raw$hourly$pm2_5)
)
```

Plot it!

```{r}
ggplot(aq_df, aes(x = time, y = pm2_5)) +
  geom_line(colour = "steelblue") +
  geom_hline(yintercept = 12, linetype = "dashed", colour = "orange") +
  annotate("text", x = min(aq_df$time), y = 13.5,
           label = "EPA annual standard (12 µg/m³)",
           hjust = 0, size = 3, colour = "orange") +
  labs(
    title = "PM2.5 Air Quality — Santa Barbara",
    x     = "Time",
    y     = "PM2.5 (µg/m³)"
  ) +
  theme_minimal()
```

Modify the request

Check what are the required parameters, eg lat and long needed for this data set.

```{r}
# Build new request 

req2 <- request("https://air-quality-api.open-meteo.com/v1/air-quality") %>%
  req_url_query(
    latitude = 34.42,
    longitude = -119.70,
    hourly = "pm2_5, ozone",
    past_days = 3,
    forecast_days = 6 # add forecast days 
  )

# Send the request
resp2 <- req %>% req_perform()

# what is this 
attributes(resp2)

# Convert  to an R list
aq2_raw <- resp %>% resp_body_json(simplifyVector = TRUE)

# What did we get back?
names(aq2_raw)

names(aq2_raw$hourly)

# Step 2 — What's inside a key?
names(aq_raw$hourly)          # compact overview of structure
# compare with previous call - longer time
# because we added forecast days
length(aq_raw$hourly$pm2_5)


```

Quiz get mean pm10 and mean pm2_5

```{r}

# Build the request 
req3 <- request ("https://air-quality-api.open-meteo.com/v1/air-quality") %>% 
  req_url_query(
    latitude = 34.42, 
    longitude = -119.70, 
    hourly = "pm2_5,pm10", 
    past_days = 3 # last 3 days of data 
  )

# send the quest and parse the response

resp3 <- req3 %>% req_perform()

# Convert to an R list 
aq_raw3 <- resp3 %>% resp_body_json(simplifyVector = TRUE) #keeps NAs

# Show what we got back 
names(aq_raw3)
str(aq_raw3) #compact overview of structure 
length(aq_raw3$pm2_5) #number of observations


```

Make the data frame

```{r}

# Extract what you need for the df 
aq_df3 <- data.frame(
  time = as.POSIXct(unlist(aq_raw$hourly$time), format = "%Y-%m-%dT%H:%M"), # shows how to interpret JSON date format 
  pm2_5 = unlist(aq_raw3$hourly$pm2_5),
  pm10 = unlist(aq_raw3$hourly$pm10)
)

str(aq_df3)

```

Calculate the mean

```{r}

mean_pm10 <- mean(aq_df3$pm10, na.rm = TRUE)
mean_pm25 <- mean(aq_df3$pm2_5, na.rm = TRUE)
  

```

Make a function with APIs

Input: latitude, longitude

Output: a tidy data frame of hourly PM2.5 readings

Parameters: number of past days

packages for function error messages: stop, warning

```{r}
get_pm25 <- function(lat, lon, past_days = 7) {
  
  # Error checking
  if (lat < -90 | lat > 90) {
    return("Latitude must be between -90 and 90.")
  }
  if (lon < -180 | lon > 180) {
    return("Longitude must be between -180 and 180.")
  }                                                    # ← added this
  
  req <- request("https://air-quality-api.open-meteo.com/v1/air-quality") |>
    req_url_query(
      latitude  = lat,
      longitude = lon,
      hourly    = "pm2_5",
      past_days = past_days
    )
  resp <- req |> req_perform()
  raw  <- resp |> resp_body_json(simplifyVector = TRUE)
  data.frame(
    lat   = lat,
    lon   = lon,
    time  = as.POSIXct(unlist(raw$hourly$time), format = "%Y-%m-%dT%H:%M"),
    pm2_5 = unlist(raw$hourly$pm2_5)
  )
}

```

Use the function with multiple locations

```{r}
# Define several CA coastal cities
locations <- data.frame(
  city = c("Santa Barbara", "Ventura", "San Luis Obispo"),
  lat  = c(34.42,           34.27,     35.28),
  lon  = c(-119.70,         -119.29,   -120.66)
)

# Use for loop to call our function for each row

for (i in 1:nrow(locations)) {
  city <- locations$city[i]
  lat  <- locations$lat[i]
  lon  <- locations$lon[i]
  
  aq_df <- get_pm25(lat, lon, past_days = 3)
  aq_df$city <- city
  
  if (i == 1) {
    aq_all <- aq_df
  } else {
    aq_all <- bind_rows(aq_all, aq_df)
  }
}


# Daily averages by city
aq_daily <- aq_all %>%
  mutate(date = as.Date(time)) %>%
  group_by(city, date) %>%
  summarize(mean_pm2_5 = mean(pm2_5, na.rm = TRUE), .groups = "drop")

ggplot(aq_daily, aes(x = date, y = mean_pm2_5, colour = city)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(
    title  = "Daily Mean PM2.5 — Central Coast Cities",
    x      = "Date",
    y      = "PM2.5 (µg/m³)",
    colour = "City"
  ) +
  theme_minimal()
```

```{r}

get_pm25(100, -119) # error message test
```
